### 2014-11-06 fundamentals/intro.tex
more edits

Historically, the national laboratories have driven development and regulated 
the use of their own tools such as \gls{VISION}\cite{jacobson_verifiable_2010}, 
\gls{DANESS}\cite{van_den_durpel_daness_2009}, 
\gls{DYMOND}\cite{yacout_modeling_2005}, and 
\gls{NFCSim}\cite{schneider_nfcsim:_2005}. 
Outside of the national laboratories, researchers create new nuclear fuel cycle 
simulation tools when existing tools are not available or not sufficiently 
general to calculate their metrics of interest with their method of choice.  
With limited access to the national laboratory tools and a need to customize 
them for research purposes, universities and private industry researchers 
``reinvent the wheel'' by developing tools of their own from scratch and 
tailored to their own needs, for example \gls{CAFCA}\cite{guerin_impact_2009} and 
\gls{COSI}\cite{boucher_cosi_2005,boucher_cosi:_2006,meyer_new_2009,coquelet-pascal_comparison_2011} respectively.

This trend has resulted in a proliferation of specialized
fuel cycle simulators for the calculation of fuel cycle metrics.  Accordingly, methods 
for calculating those metrics vary among simulators. Some model the 
system of facilities, economics, and materials at static equilibrium. Meanwhile 
other simulators capture its dynamics.  
Similarly, while some simulators discretely model material and facilities, 
others treat materials in streams and facilities in fleets, in an aggregated 
fashion. Depending on their intended use, simulators often overspecialize, 
modeling a single aspect of the fuel cycle in great detail while neglecting 
others. For example, a simulator created for policy modeling might have excellent 
capability in economics while material isotopics and transmutation detail are 
neglected.

The \Cyclus simulator emerged from a line of tools seeking to break this practice.  
A \gls{LDRD} project that began at \gls{INL}, 
\gls{GENIUSv1}\cite{dunzik-gougar_global_2007,jain_transitioning_2006} sought 
to provide generic regional capability.  Based on lessons learned from 
\gls{GENIUSv1}, the 
\gls{GENIUSv2}\cite{oliver_studying_2009,huff_geniusv2_2009} code sought to 
provide more generality and an extensible interface to facilitate 
collaboration.  The \Cyclus project, in turn, improved upon the \gls{GENIUSv2} 
effort by implementing increased modularity and encapsulation.  The result is  
a dynamic simulator that treats both materials and facilities discretely, with 
an architecture that permits simulations with multiple and variable levels of 
fidelity. The simulator's agent-based framework for tracking of 
transformation and trade of resources between facilities within institutions 
and regions greatly improves upon the existing fuel cycle simulators.

@@

In addition to providing the fundamental features expected in fuel cycle 
simulators, the \Cyclus paradigm enables essential capabilities that previous 
simulators lack. First, its \emph{modular and open architecture}  facilitates 
targeted contribution and collaboration within the nuclear fuel cycle analysis 
community.  Additionally, \emph{agent interchangeability} facilitates direct 
comparison of modeling methodologies. Finally, the flexibility and generality 
of its \emph{\gls{ABM}} paradigm and \emph{discrete, object-oriented approach} 
emphasizes applicability of the \Cyclus tool across a broader range of 
fidelities, scales, and applications than other simulators.

\Cyclus separates the problem of modeling a
physics-dependent supply chain into a simulation kernel and archetypes that
interact with the kernel during a simulation. The kernel is responsible for the
deployment and interaction logic of entities in the simulation. Physics
calculations are constrained to individual archetypes. Accordingly, specialists
in a given field can utilize their time and resources in modeling a specific
process associated with the nuclear fuel cycle (e.g., reprocessing and advanced
fuel fabrication), rather than also modeling the entirety of the cycle.

Furthermore, modeling the evolution of a physics-dependent, international nuclear fuel supply
chain is a multiscale problem, but existing tools are not sufficiently 
multiscale. Tools have either focused on macro effects, e.g., the fleet-level
stocks and flows of commodities, or micro effects, e.g., the used fuel
composition of fast reactor fuel. Each focus has driven the
development of these tools, rendering the task of answering questions between
the macro and micro levels challenging within a single tool. 
However, an extensible architecture and discrete object tracking allow 
the creation and interchangeability of custom \Cyclus archetypes at any level of fidelity.

@@

The architecture within the \Cyclus tool allows developers 
to define nuclear fuel cycle agents and processes independent of the simulation 
logic. This `encapsulated,' `plug-in'
design choice provides two major benefits.

First, analysts across research institutions can take advantage of the 
simulation logic \gls{API} and archetype ecosystem when they apply \Cyclus to 
their specific problem. This extensibility, magnified by an open source 
development process, frees the community of users to contribute desired 
features, independent from the team of \Cyclus core developers. This 
mend-and-extend paradigm, common in modern software development, reduces 
maintenance overhead and improves code robustness \cite{citation_needed}. 
The researcher can therefore focus on creating or customizing nuclear 
facility, institution, resource, and toolkit models within their specific area 
of technical expertise.

% we could use a specific example of Cosi's equivalence method below
Second, because \Cyclus uses a modular archetype approach, comparing two 
archetypes is straightforward. Consider a case in which an analyst would like 
to compare the effect of using different models to determine the input fuel 
composition for fast reactors. A variety of fuel fabrication archetypes can be 
developed and interchanged to analyze the effect of each on the 
simulation output.

There are a number of types of users and developers who will 
interact with the \Cyclus framework. National laboratory researchers,
academic researchers, private-industry economic forecasters, and even students 
in engineering courses may benefit from the ease of development and
customization provided by the modular architecture of \Cyclus.

@@

\Cyclus implements an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and it takes advantage of modern computing resources.  
Furthermore, agent-based modeling provides an analyst a more natural 
\Cyclus utilizes an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and takes advantage of modern computing resources.  
Furthermore, \gls{ABM} provides an analyst a more natural 
perspective from which to model a given problem. In the nuclear fuel cycle 
context for example, an analyst can design an agent with reactor-related logic 
entirely independent from a separate agent with fuel fabrication-related 
logic. Defining the behavior of both agents according to the 
\gls{API} contract is sufficient for them to interact with one another as 
bona fide actors in the simulation.  The two archetype libraries, therefore, need 
not have any shared knowledge of one another for them to be building blocks in 
the same simulation.

System dynamics is a popular approach for modeling nuclear fuel cycles 
\cite{jacobson_vision_2009,van_den_durpel_daness_2009,guerin_impact_2009,guerin_benchmark_2009}. However, it has been shown in systems analysis literature
\cite{macal_agent-based_2010} that, formally, the
set of system dynamics models is a strict subset of agent-based models.
That is, any system dynamics model can be translated
into an agent-based model. Furthermore, agent-based
techniques can provide a richer model than can system dynamics. To enable the 
broadest range of simulations in a generic fashion, \gls{ABM} is therefore 
superior to system dynamics.

@@

This generality specifically permits many use cases that cannot be achieved with 
fleet-based models. That is, some questions require resolution at the level of 
individual facilities and materials.  Since fleet-based, 
lumped-material models do not distinguish between discrete facility entities or 
materials, many detailed performance metrics cannot be captured with 
those tools.

One example use case seeks to capture system vulnerability to nefarious 
material diversion. This type of analysis requires discrete simulation of a 
target facility and the individual materials modified within it. 
Provenance and trade-history of distinct materials is the fundamental 
information unit in such studies. 
Material risk analysis, therefore, demands discretely modeled objects like those 
in the \Cyclus simulator.

>>>>

Historically, the national laboratories have driven development and regulated 
the use of their own tools: \gls{VISION}\cite{jacobson_verifiable_2010}, 
\gls{DANESS}\cite{van_den_durpel_daness_2009}, 
\gls{DYMOND}\cite{yacout_modeling_2005}, and 
\gls{NFCSim}\cite{schneider_nfcsim:_2005}. 
Outside the national laboratories, researchers have created new nuclear fuel cycle 
simulation tools when existing tools were not available or not sufficiently 
general to calculate their metrics of interest with their method of choice.  
With limited access to the national laboratory tools and a need to customize 
them for research purposes, universities and private industry researchers have 
``reinvented the wheel'' by developing tools of their own from scratch and 
tailored to their own needs, for example \gls{CAFCA}\cite{guerin_impact_2009} and 
\gls{COSI}\cite{boucher_cosi_2005,boucher_cosi:_2006,meyer_new_2009,coquelet-pascal_comparison_2011}.

This trend has resulted in a proliferation of specialized
fuel cycle simulators for the calculation of fuel cycle metrics.  Accordingly, methods 
for calculating those metrics vary among simulators. Some model the 
system of facilities, economics, and materials in static equilibrium, while 
other simulators capture the dynamics of the system.  
Similarly, while some simulators discretely model material and facilities, 
others treat materials as streams and facilities as fleets, in an aggregated 
fashion. Simulators often overspecialize, 
modeling a single aspect of the fuel cycle in great detail while neglecting 
others. For example, a simulator created for policy modeling might have excellent 
capability in economics while material isotopics and transmutation detail are 
neglected.

The \Cyclus simulator emerged from a line of tools seeking to break this practice.  
Its precurser, \gls{GENIUSv1}\cite{dunzik-gougar_global_2007,jain_transitioning_2006}, was begun at \gls{LDRD}, began at \gls{INL} and sought 
to provide generic regional capability.  Based on lessons learned from 
\gls{GENIUSv1}, the 
\gls{GENIUSv2}\cite{oliver_studying_2009,huff_geniusv2_2009} code sought to 
provide more generality and an extensible interface to facilitate 
collaboration.  The \Cyclus project then improved upon the \gls{GENIUSv2} 
effort by implementing increased modularity and encapsulation.  The result is  
a dynamic simulator that treats both materials and facilities discretely, with 
an architecture that permits simulations with multiple and variable levels of 
fidelity. The simulator's agent-based framework for tracking the 
transformation and trade of resources between facilities within institutions 
and regions greatly improves upon the existing fuel cycle simulators.

@@

In addition to providing the fundamental features expected in fuel cycle 
simulators, the \Cyclus paradigm enables essential capabilities that previous 
simulators lack. Critically, its \emph{modular and open architecture}  facilitates 
targeted contribution and collaboration within the nuclear fuel cycle analysis 
community.  Additionally, \emph{agent interchangeability} facilitates direct 
comparison of modeling methodologies. Finally, the flexibility and generality 
of its \emph{\gls{ABM}} paradigm and \emph{discrete, object-oriented approach} 
enables \Cyclus to be applicable to a broader range of 
fidelities, scales, and applications than to other simulators.

\Cyclus separates the problem of modeling a
physics-dependent supply chain into two distinct components: a simulation kernel and archetypes that
interact with the kernel. The kernel is responsible for the
deployment and interaction logic of entities in the simulation. The archetypes represent 
individual types of facilities and capture their unique physics.  Accordingly, specialists
in a given field can utilize their time and resources in modeling a specific
process associated with the nuclear fuel cycle (e.g., reprocessing and advanced
fuel fabrication), without having to model the entirety of the cycle.

Furthermore, modeling the evolution of a physics-dependent, international nuclear fuel supply
chain is a multiscale problem, and existing tools are not sufficiently 
multiscale. Tools have either focused on macro effects, e.g., the fleet-level
stocks and flows of commodities, or micro effects, e.g., the used-fuel
composition of fast reactor fuel. Each focus has driven the
development of specialized tools, rendering the task of answering questions between
the macro and micro levels challenging within a single tool. 
In constrast, the extensible architecture and discrete object tracking of \Cyclus allow
the creation and interchangeability of custom archetypes at any level of fidelity.

@@

The architecture of \Cyclus allows developers 
to define nuclear fuel cycle agents and processes independent of the simulation 
logic. This encapsulated `plug-in'
design choice provides two major benefits. First, analysts across research institutions can take advantage of the 
simulation logic \gls{API} and archetype ecosystem when they apply \Cyclus to 
their specific problem. This 
mend-and-extend paradigm, common in modern software development, reduces 
maintenance overhead and improves code robustness \cite{citation_needed}. 
The researcher can focus on creating or customizing nuclear 
facility, institution, resource, and toolkit models within their specific area 
of technical expertise. This extensibility, magnified by an open source 
development process, frees the community of users to contribute desired 
features, independent from the team of \Cyclus core developers. 
% we could use a specific example of Cosi's equivalence method below
Second, because \Cyclus uses a modular archetype approach, comparing two 
archetypes is straightforward. Consider a case in which an analyst would like 
to compare the effect of using different models to determine the input fuel 
composition for fast reactors. A variety of fuel fabrication archetypes can be 
developed and interchanged while keeping the rest of the fuel cycle fixed
 to compare fuel fabrication designs.

@@

\Cyclus implements an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and it incorporates modern computing resources.  
Furthermore, \gls{ABM} provides a natural 
perspective from which to model a given problem. In the nuclear fuel cycle 
context, for example, an analyst can design an agent with reactor-related logic 
that is entirely independent from an agent with fuel fabrication-related 
logic. Defining the behavior of both agents according to the 
\gls{API} contract is sufficient for them to interact with one another as 
bona fide actors in the simulation.  The two archetype libraries can be used in the same
simulation without any shared knowledge.

Furthermore, the \gls{ABM} paradigm is easily comparable to system dynamics models.
System dynamics is a popular approach for modeling nuclear fuel cycles 
\cite{jacobson_vision_2009,van_den_durpel_daness_2009,guerin_impact_2009,guerin_benchmark_2009}. 
Formally however, system dynamics models are simply a strict subset of agent-based models
\cite{macal_agent-based_2010}.
That is, any system dynamics model can be translated
into an agent-based model. Furthermore, because agent-based
techniques can provide a richer model than can system dynamics, they enable the 
broadest range of simulations in a generic fashion.

@@

Although some questions require resolution at the level of 
individual facilities and materials, 
Fleet-based, lumped-material models do not distinguish between discrete
facility entities or materials.  As a result, detailed performance 
metrics cannot be captured with previously existing tools.
The generality of \Cyclus therefore enables many use cases 
that cannot be achieved with fleet-based models.

One such example is the use case that seeks to capture system vulnerability to nefarious 
material diversion. Provenance and trade-history of distinct materials is the fundamental 
information unit in such studies, and so this type of analysis requires
 discrete simulation of a 
target facility and the individual materials modified within it. 
Provenance and trade-history of distinct materials is the fundamental 
information unit in such studies. 
Material risk analysis, therefore, demands discretely modeled objects like those 
in the \Cyclus simulator.

There are a number of types of users and developers who will 
interact with the \Cyclus framework. National laboratory researchers,
academic researchers, private-industry economic forecasters, and even students 
in engineering courses may benefit from the ease of development and
customization provided by the discrete, modular, agent-based architecture of \Cyclus.
